---
title: "PSIR Graduate Methods Problem Set 1"
author: "Hoik Jung 2026-27005"
date: "`r format(Sys.time(), '%B %d, %Y')`"
header_includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{amsthm}
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
---

# Instructions
* All answers must be typed up and submitted electronically as a PDF file, either via LaTeX or R Markdown. Submit the corresponding .Rmd or .R file as well.
* Most problem sets should be completed within the individual assigned groups. Collaboration is not allowed for the problem sets between set groups/individuals.
* Substitute "author" with your own names for the submission.
* Problem set 1 will be completed in groups of 4--5.

\newpage

# The Average Treatment Effect and the Independence Assumption: Applying it to Research

This part should have every member's individual answers assembled here as subsections. I know you already submitted the abstract/intro, but let's do the following again to clarify the RQ and the estimation strategy:

1. What is the *causal* research question that you want to tackle in the final paper? State the question clearly. This does not mean that you have a clean causal setup; even when checking for correlations, the underlying questions are typically causal.
2. How would you operationalize and measure your outcome variable?
3. How would you operationalize and measure your main explanatory variable?
4. Consider the *average treatment effect (ATE)* under the potential outcomes framework, given binary treatment. If your explanatory variable is continuous or discrete, try creating a binary explanatory variable based on it, such as dichotomizing it. Consider the difference-in-means estimator (subscripts discarded for brevity):

$$ATE = \mathop{\mathbb{E}}[Y \mid T=1] - \mathop{\mathbb{E}}[Y \mid T=0]$$

In your research question, is this a valid estimator of the average treatment effect? That is to say, is the independence assumption $Y_i(1), Y_i(0) \perp T_i$ satisfied? At this point, do not consider controlling for other variables. We will return to that possibility later in the course.


-------------------------------------------
 Answer 
 
#1. Causal research question
Does owning stock market assets (individual stocks or mutual funds) causally shift Americans' economic and trade policy attitudes toward a more market-friendly, individualist position — or do stockholders simply happen to hold such views for other reasons?

#2. Outcome variable
Economic/trade policy attitude, operationalized two ways from ANES: (a) a 7-point free-trade-agreement favor/oppose scale (V162176x/V202361x/V242346x), rescaled to [0,1]; (b) a 7-point "government should guarantee jobs" economic-liberalism scale (V161189/V201255/V241252), where higher values indicate preference for individual responsibility over government intervention.

#3. Main explanatory variable 
A binary indicator for stock ownership, from the ANES item "Do you personally, or jointly with a spouse, have any money invested in the stock market right now — either in an individual stock or in a mutual fund?" ($T_i=1$ if yes, $0$ if no).

#4. Is the difference-in-means a valid ATE estimator?
No. $T_i$ is self-selected, not randomly assigned, so independence $Y_i(1), Y_i(0) \perp T_i$ almost certainly fails. Stockholders differ systematically from non-owners on income, education, and homeownership (I verified this directly with a balance table: standardized mean differences of 0.73–0.94 on income/education), and plausibly on unobserved traits like risk tolerance — all of which independently predict trade/economic attitudes. A naive difference-in-means would therefore conflate the true treatment effect with selection bias.
 
 
----------------------------------------------

\newpage

# The Average Treatment Effect in a Randomized Experiment

White evangelical Republicans often hold two identities that point in different directions on refugee policy. Republican leaders have generally favored restrictions, while Christian teachings can emphasize care for refugees and other vulnerable strangers. DeMora et al. (2024) ask whether a message framed around religious values can increase support for refugees among self-identified white evangelical Republicans.

This exercise is based on: DeMora, Stephanie L., Jennifer L. Merolla, Brian Newman, and Elizabeth J. Zechmeister. 2024. "[Jesus Was a Refugee: Religious Values Framing Can Increase Support for Refugees Among White Evangelical Republicans.](https://doi.org/10.1007/s11109-024-09912-2)" *Political Behavior* 46: 2145--2168. The authors' replication materials are available from the [Harvard Dataverse](https://doi.org/10.7910/DVN/QYZQ4Y).

Read the sections titled "An Experiment to Assess the Religious Values Frame" and "Results of the Religious Values Framing Experiment" carefully. The researchers randomly assigned respondents to a control condition, a religious-values message without a source cue, or the same message accompanied by an evangelical source cue. The control group received neither the message nor its two follow-up questions.

The supplied file, `data/demora_2024.csv`, is a simplified teaching version of the authors' cleaned YouGov sample. Each row describes one respondent. Higher values on every outcome indicate more favorable attitudes toward refugees. Do not use `sample_weight` unless a question specifically asks you to do so.

--------------------------------------------------------------------------------
 Name                       Description
 -------------------------- -----------------------------------------------------
 `respondent_id`            Anonymous row identifier

 `treatment`                Control, religious-values message, or religious-values message with a source cue

 `refugee_thermometer`      Feeling toward refugees, from 0 to 100

 `resettlement_support`     Support for refugee resettlement, from 0 to 1

 `school_support`           Support for refugee children attending public schools, from 1 to 5

 `benefits_support`         Support for refugees receiving unemployment benefits, from 1 to 5

 `age`                      Respondent's age in 2020

 `gender`                   Respondent's gender, recorded as `Male` or `Female`

 `education`                Respondent's highest level of education

 `ideology`                 Ideology from 1, very liberal, to 5, very conservative

 `sample_weight`            YouGov weight for the target population
--------------------------------------------------------------------------------

Load the data before beginning the exercise. Beware of the relative paths; you may put the data where it suits you and edit the file path used for `read.csv`.

```{r, message=FALSE, warning=FALSE}
demora <- read.csv("data/demora_2024.csv")

demora$treatment <- factor(
  demora$treatment,
  levels = c(
    "Control",
    "Religious values message",
    "Religious values + source cue"
  )
)

summary(demora)
```

## Question 1: Research Design

1. State the paper's causal research question in your own words.
2. What are the three experimental conditions? What features distinguish them?
3. What is the unit of random assignment? What is the unit of observation in the supplied data?
4. Focus first on the religious-values message without a source cue versus the control condition. Define $T_i$, $Y_i(1)$, $Y_i(0)$, and the average treatment effect when `refugee_thermometer` is the outcome.
5. Why can we never directly observe $Y_i(1)-Y_i(0)$ for any respondent?

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]

#1.
RQ: Can religious vlaues framing move partisans away from the party line on policy attitudes? 

#2. 
The experiment contains three randomly assigned conditions ; Two Treatment groups, and one control group. The Religious Values condition introduces a moral frame emphasizing core Christian teachings. The Religious values + Source cue condition contains the exact same treatment with Religious value group but adds an explicit elite attribution stating. While control group 
receives no text or message, with respondents moving directly to post-treatment survey question to establish a baseline measure of unprompted political attitudes. 

#3.
Unit of Random Assignment : The individual survey respondent (participant). Each individual participant was randomly assigned to one of the three experimental conditions. 

Unit of Observation: The individual respondent (individual level). Each observation/row in the dataset represents a single survey participant, containing their demographic traits, assigned treatment condition, identity strength scores, and post-treatment responses. 


#4.
Ti is a binary indicator variable representing whether the respondent is assigned to receive the generic religious value messages or not. Ti=1 if respondent I was assigned to the Religious Values condition. Ti=0 if respondent I was assigned to the control condition. 

Yi(1) is the potential feeling thermometer score toward refugees that respondent I would express if exposed to the Religious Values Treatment (Ti=1)

Yi(0) is the potential feeling thermometer score toward refugees that responded I would express if assigned to the control condition (Ti=0) 

The Average Treatment Effect (ATE) is the expected population difference between the two potential outcomes across all individuals in the target population . 
ATE = E[Yi(1)-Yi(0)] = E[Yi(1)] - E[Yi(0)]
The Average Treatment Effect (ATE) measures the expected causal impact across the entire target population rather than just the treated group. Because random assignment makes treatment status statistically independent of potential outcomes, the observed difference between treatment and control group averages provides an unbiased estimate of the overall ATE.


#5.
We cannot directly observe Yi(1) - Yi(0) because each respondent can be assigned to only one treatment condition, so we can not observe the both potential outcomes of Yi(1) and Yi(0). We call this ‘the fundamental problem of causal inference’. Random assignment allows us to overcome this fundamental problem of causal inference at the group level by using control group to estimate the average counterfactual outcome for the treated group. 

--------------------------------------------------------------------------------

## Question 2: Exploring the Experiment

1. How many observations and variables are in the supplied data?
2. How many respondents are in each experimental condition?
3. Which variables contain missing values? Show how you checked.
4. Report the mean and standard deviation of age and the proportion of respondents recorded as female.
5. Describe the distributions and scales of the four outcome variables. Explain why their numerical values should not be compared without attending to their scales.

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]

```{r}
#1  
 dim(demora) #dim
 nrow(demora)   # observations
 ncol(demora)   # variables
 
#2. 
 table(demora$treatment)
 
#3. 
 colSums(is.na(demora)) #refugee_theormometer: 25, ideology:7
 
#4.
#age 
 mean(demora$age, na.rm = TRUE) #54.96481
 sd(demora$age, na.rm = TRUE) #17.77131
#female 
 mean(demora$gender == "Female", na.rm = TRUE) #mean=0.527
 prop.table(table(demora$gender)) #0.527
 
 #5.
 # Define outcome variables
outcomes <- c(
  "refugee_thermometer",
  "resettlement_support",
  "school_support",
  "benefits_support"
)

# Summary statistics
sapply(demora[outcomes], function(x) {
  c(
    Min  = min(x, na.rm = TRUE),
    Max  = max(x, na.rm = TRUE),
    Mean = mean(x, na.rm = TRUE),
    SD   = sd(x, na.rm = TRUE)
  )
})

lapply(demora[outcomes], table, useNA = "ifany")

#their raw numerical values should not be directly compared because the same numerical difference represents different amounts of change across these scales.
 
```


--------------------------------------------------------------------------------

## Question 3: Estimating Average Treatment Effects

1. Calculate the unweighted mean of each outcome separately for the three experimental conditions.
2. For each outcome, estimate the effect of the religious-values message relative to the control using a difference in means.
3. For each outcome, estimate the effect of the religious-values message with a source cue relative to the control using a difference in means.
4. Interpret every estimate in the units of its outcome. For example, distinguish points on the 0--100 feeling thermometer from changes on the 0--1 resettlement scale.
5. Create one clearly labeled figure comparing mean `refugee_thermometer` scores across the three conditions. Preserve the experimental ordering shown in the variable table.
6. Based on these estimates, does the source cue appear to strengthen the message? Explain what comparison answers this question.

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]

```{r}
## [PUT YOUR CODE HERE]
#1. 
aggregate(
  demora[outcomes],
  by = list(treatment = demora$treatment),
  FUN = mean,
  na.rm = TRUE
)

library(dplyr)  

demora %>%
  group_by(treatment) %>%
  summarise(
    across(
      all_of(outcomes),
      ~ mean(.x, na.rm = TRUE)
    )
  )

means <- sapply(demora[outcomes], function(x) {
  tapply(x, demora$treatment, mean, na.rm = TRUE)
})
print(means)


#2. 
sapply(outcomes, function(y) {
  mean(demora[demora$treatment == "Religious values message", y],
       na.rm = TRUE) -
  mean(demora[demora$treatment == "Control", y],
       na.rm = TRUE)
})

#3. 
sapply(outcomes, function(y) {
  mean(demora[demora$treatment == "Religious values + source cue", y],
       na.rm = TRUE) -
    mean(demora[demora$treatment == "Control", y],
         na.rm = TRUE)
})

#4. #explanation below. 
# Religious-values message vs. Control
message_effect <- sapply(outcomes, function(y) {
  mean(demora[demora$treatment == "Religious values message", y],
       na.rm = TRUE) -
    mean(demora[demora$treatment == "Control", y],
         na.rm = TRUE)
})

# Religious-values message + source cue vs. Control
source_effect <- sapply(outcomes, function(y) {
  mean(demora[demora$treatment == "Religious values + source cue", y],
       na.rm = TRUE) -
    mean(demora[demora$treatment == "Control", y],
         na.rm = TRUE)
})

effects <- data.frame(
  Outcome = outcomes,
  Message_vs_Control = message_effect,
  SourceCue_vs_Control = source_effect
)

effects


#5. #explanation below 
thermo_means <- demora %>%
  group_by(treatment) %>%
  summarise(
    mean_thermometer = mean(refugee_thermometer, na.rm = TRUE),
    sd_thermometer = sd(refugee_thermometer, na.rm = TRUE),
    n = sum(!is.na(refugee_thermometer)),
    se = sd_thermometer / sqrt(n),
    lower = mean_thermometer - 1.96 * se,
    upper = mean_thermometer + 1.96 * se
  )

library(ggplot2)
ggplot(
  thermo_means,
  aes(x = treatment, y = mean_thermometer)
) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    width = 0.12
  ) +
  geom_point(size = 3) +
  labs(
    title = "Mean Refugee Thermometer Score by Experimental Condition",
    x = "Experimental Condition",
    y = "Mean Refugee Thermometer Score (0–100)"
  ) +
  ylim(0, 100) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 20, hjust = 1)
  )
```
#4.
Relative to the control, the religious-values message increased the refugee thermometer by 9.36 points on the 0–100 scale and resettlement support by 0.018 units on the 0–1 scale, while decreasing school support by 0.018 points and benefits support by 0.141 points on their 1–5 scales. With the source cue, the corresponding changes were +5.32 points, +0.019 units, +0.006 points, and −0.137 points, respectively.


#6.
No. Based on the point estimates, the source cue does not appear to strengthen the message. The relevant comparison is the message-plus-source-cue condition versus the message-only condition. The estimated effects relative to control are 5.32 and 9.36 points, respectively, so the direct difference is \(5.32-9.36=-4.04\) points on the 0–100 refugee thermometer scale.

------------------------------------------------------------------------------------------

## Question 4: Randomization and Independence

1. Compare age, gender, education, and ideology across the three conditions. Present a readable balance table. You will need to explain how you summarized categorical variables.
2. Are the groups exactly identical on these observed pretreatment characteristics? Should random assignment make them exactly identical in this realized sample?
3. What does random assignment imply about the relationship between treatment assignment and both observed and unobserved pretreatment characteristics across repeated assignments?
4. Based on the research design, is $Y_i(1),Y_i(0) \perp T_i$ plausible for the religious-values-message-versus-control comparison? Explain why the design, rather than the observed balance table alone, is the basis for your answer.
5. Suppose one balance difference is statistically significant at the 0.05 level after examining twenty pretreatment variables. Would this fact alone invalidate the experiment? Explain.

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]

```{r}
#1. 
# Load packages
library(dplyr)
library(gtsummary)

# Check variables
str(demora[c("age", "gender", "education", "ideology")])

table(demora$gender, useNA = "ifany")
table(demora$education, useNA = "ifany")
table(demora$ideology, useNA = "ifany")


# Treat ideology (1-5 scale) as a categorical/ordinal variable
demora <- demora %>%
  mutate(
    ideology = factor(
      ideology,
      levels = 1:5,
      ordered = TRUE
    )
  )


# Create balance table
balance_table <- demora %>%
  select(treatment, age, gender, education, ideology) %>%
  tbl_summary(
    by = treatment,
    
    # Explicitly specify variable types
    type = list(
      age ~ "continuous",
      gender ~ "categorical",
      education ~ "categorical",
      ideology ~ "categorical"
    ),
    
    # How each variable is summarized
    statistic = list(
      age ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    
    # Number of decimal places
    digits = list(
      age ~ 1,
      all_categorical() ~ c(0, 1)
    ),
    
    missing = "ifany"
  ) %>%
  modify_header(
    label ~ "**Variable**"
  ) %>%
  bold_labels()

# Display table
balance_table

#2.

# Age
demora %>%
  group_by(treatment) %>%
  summarise(
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE)
  )

# Gender
prop.table(
  table(demora$treatment, demora$gender),
  margin = 1
)

# Education
prop.table(
  table(demora$treatment, demora$education),
  margin = 1
)

# Ideology
prop.table(
  table(demora$treatment, demora$ideology),
  margin = 1
)

 
```
#1. 
Categorical variables (gender, education, and ideology) were summarized using the proportion of respondents in each category within each experimental condition.

#2.
No. The groups are not exactly identical on the observed pretreatment characteristics, as the balance table shows small differences in age and in the distributions of gender, education, and ideology across conditions. Random assignment does not guarantee exact balance in a particular realized sample. Rather, it produces balance in expectation across repeated random assignments, so some chance differences between groups are expected.
#3.
Across repeated random assignments, treatment assignment is independent of both observed and unobserved pretreatment characteristics. Thus, treatment groups are balanced on these characteristics in expectation, although chance imbalances may occur in any particular realized sample. 
#4.
Balance is a diagnostic check; randomization is the identification strategy. The research design, especially random assignment, is the basis for causal inference because it makes treatment assignment independent of potential outcomes by construction. An observed balance table only shows that the treatment and control groups happen to be similar on measured covariates in the realized sample. It cannot establish balance on unobserved characteristics. Therefore, balance tables are useful as diagnostic checks, but randomization—not observed balance—is what justifies causal identification.
#5. 
No. A single statistically significant balance difference would not by itself invalidate the experiment. When twenty pretreatment variables are tested at the 0.05 significance level, some significant differences can arise by chance even under proper random assignment. In fact, the probability of observing at least one significant result among twenty independent tests is approximately \(1-0.95^{20}\approx 0.64\). Therefore, an isolated significant imbalance is not sufficient evidence that randomization failed. The validity of the experiment should primarily be assessed based on the random assignment procedure rather than whether every observed covariate is statistically balanced.


---------------------------------------------------------------------------------------
## Question 5: Measurement, Assumptions, and Generalization

1. The `resettlement_support` outcome combines support for resettlement in the respondent's local community and in the United States. Discuss one advantage and one disadvantage of combining the two items.
2. The authors note that the treatment technically consists of a message plus two questions that ask respondents to reflect on it. What complication does this create when naming the causal treatment?
3. About 77 percent of treated respondents reported having heard the message previously. Does prior exposure violate random assignment? How might it affect the treatment effect estimated in this experiment?
4. State the Stable Unit Treatment Value Assumption (SUTVA) in this study. Give one plausible example of interference or treatment variation.
5. The analytic sample consists of self-identified white evangelical Republicans who passed the authors' prespecified reading-speed screen. Identify two limits this places on external validity.
6. The outcomes are survey responses measured shortly after treatment. What can and cannot be concluded about lasting changes in attitudes or political behavior?

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]
1. Advantage: Combining the two items can reduce measurement noise and capture a broader, more reliable measure of general support for refugee resettlement.
Disadvantage: It may hide meaningful differences between support for local resettlement and support for national resettlement, especially if respondents support refugees nationally but not in their own community.

2. #page 9 참고 (수정) 

3. Prior exposure does not violate random assignment, because randomization should balance prior exposure across treatment conditions on average. However, prior exposure may reduce the novelty or marginal impact of the treatment, potentially attenuating the estimated treatment effect. This does not necessarily create bias in the experimental comparison, but it may change the estimand: the experiment largely estimates the effect of additional exposure among a population already familiar with the message. Controlling for prior exposure may improve precision or allow analysis of heterogeneous treatment effects.

4. The SUTVA assumption requires that the treatment be consistently defined across participants and that the treatment assigned to one participant does not affect the outcomes of other participants. SUTVA may therefore be violated if, during the experiment, participants share treatment-related information with members of their church community or are exposed to information about the treatment through other participants.

5. External validity refers to whether the findings of an experiment can be generalized to other settings or contexts when the study is replicated under different conditions.
For this experiment to have greater external validity, we would need to examine whether voters with different social identities respond similarly to the same mechanism when considering specific political issues.

6. It is not possible to determine whether the emotional changes induced by the experimental treatment ultimately translate into actual voting behavior in elections.

--------------------------------------------------------------------------------

##Question 6: Reading the Published Results

1. Compare your unweighted estimates with Table 2 of the paper. The authors use YouGov sample weights and regression, so your estimates need not be numerically identical. Explain what each approach targets.
2. Which outcomes provide evidence that the religious-values message increased support for refugees? Which outcomes produce null results?
3. Does the paper support the hypothesis that adding an evangelical source cue makes the religious-values message more effective? Cite the relevant comparison.
4. Summarize the paper's theoretical claim and central empirical finding in no more than 150 words.
5. Propose one follow-up experiment in a different population or political context. Clearly identify the treatment, outcome, target population, and the average treatment effect of interest.

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]
#1.

Unweighted mean difference estimates a simple difference in group means, treating each respondent equally in the current analytical sample. In contrast, Table 2 of the paper estimates the treatment effect by adjusting for differences in sample composition using YouGov sample weights and regression analysis. Therefore, the numerical values do not need to be exactly the same, but we can compare the direction of the effects and the overall patterns. In the paper, the religious-values message was estimated to increase the thermometer score by approximately 11 points and resettlement support by approximately 0.05.

#2.
In the paper, a clear positive effect appeared in the refugee thermometer, and weak positive evidence also appeared in resettlement support. In contrast, no positive effects appeared in school support and unemployment/benefits support.

#3.
No. To evaluate this, the Religious Values + Source Cue condition must be directly compared with the Religious Values condition. In this direct comparison, the paper found no evidence that the source cue made the message more effective. Therefore,H2 was not supported.

#3.
The paper's theoretical argument is that framing that appeals to an individual's religious and moral values can change political attitudes even when those values conflict with partisan positions. The experimental results showed that the pro-refugee Christian values message increased favorable attitudes toward refugees and some support for refugee resettlement among White evangelical Republicans. However, it had no effect on attitudes toward government welfare support for refugees, and adding an evangelical source cue did not make the message more effective.

#4.


#5.
Indian Hindus are shown religious messages and doctrines related to peace, and their feeling thermometer toward Muslims is measured.

- Treatment: Using messages and warnings about peace from Hindu scriptures.
- Outcome: Views on the establishment of Muslim educational institutions in the local community.
- Target Population: Indian regions with a high proportion of Muslims or Indian states where Hindu-Muslim conflicts have frequently occurred in the past.
- ATE: The comparison between the treatment group that received the religious message and the control group that did not receive the message.
--------------------------------------------------------------------------------

\newpage

# Digesting Cutting Edge Methods Papers into Key Takeaways (Building an Open-Source Reviewer Agent)

See [https://github.com/sysilviakim/quant-social-science-skills/tree/main/contrib](https://github.com/sysilviakim/quant-social-science-skills/tree/main/contrib). We are going to build an open-source reviewer agent focusing on quantitative methods. 

The TA has randomly assigned two papers to you. See `EXAMPLE-atsusaka_kim_2025.md` in the GitHub repository; your task is to carefully read the papers and create something equivalent for your assigned papers.

For PS1, choose one of the two assigned papers. For that paper, submit the following two items:

1. Show me that you actually did the reading. Print out the paper, take highlights and memos, and submit the outcome in person (I don't care if the memos and highlights are messy).
2. Submit a `.md` file that summarizes the paper's main points and explains what the reviewer should check for methodological integrity.
