---
title: "PSIR Graduate Methods Problem Set 1"
author: "Hoik Jung 2026-27005"
date: "`r format(Sys.time(), '%B %d, %Y')`"
header_includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{amsthm}
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
---

# Instructions
* All answers must be typed up and submitted electronically as a PDF file, either via LaTeX or R Markdown. Submit the corresponding .Rmd or .R file as well.
* Most problem sets should be completed within the individual assigned groups. Collaboration is not allowed for the problem sets between set groups/individuals.
* Substitute "author" with your own names for the submission.
* Problem set 1 will be completed in groups of 4--5.

\newpage

# The Average Treatment Effect and the Independence Assumption: Applying it to Research

This part should have every member's individual answers assembled here as subsections. I know you already submitted the abstract/intro, but let's do the following again to clarify the RQ and the estimation strategy:

1. What is the *causal* research question that you want to tackle in the final paper? State the question clearly. This does not mean that you have a clean causal setup; even when checking for correlations, the underlying questions are typically causal.
2. How would you operationalize and measure your outcome variable?
3. How would you operationalize and measure your main explanatory variable?
4. Consider the *average treatment effect (ATE)* under the potential outcomes framework, given binary treatment. If your explanatory variable is continuous or discrete, try creating a binary explanatory variable based on it, such as dichotomizing it. Consider the difference-in-means estimator (subscripts discarded for brevity):

$$ATE = \mathop{\mathbb{E}}[Y \mid T=1] - \mathop{\mathbb{E}}[Y \mid T=0]$$

In your research question, is this a valid estimator of the average treatment effect? That is to say, is the independence assumption $Y_i(1), Y_i(0) \perp T_i$ satisfied? At this point, do not consider controlling for other variables. We will return to that possibility later in the course.

\newpage

# The Average Treatment Effect in a Randomized Experiment

White evangelical Republicans often hold two identities that point in different directions on refugee policy. Republican leaders have generally favored restrictions, while Christian teachings can emphasize care for refugees and other vulnerable strangers. DeMora et al. (2024) ask whether a message framed around religious values can increase support for refugees among self-identified white evangelical Republicans.

This exercise is based on: DeMora, Stephanie L., Jennifer L. Merolla, Brian Newman, and Elizabeth J. Zechmeister. 2024. "[Jesus Was a Refugee: Religious Values Framing Can Increase Support for Refugees Among White Evangelical Republicans.](https://doi.org/10.1007/s11109-024-09912-2)" *Political Behavior* 46: 2145--2168. The authors' replication materials are available from the [Harvard Dataverse](https://doi.org/10.7910/DVN/QYZQ4Y).

Read the sections titled "An Experiment to Assess the Religious Values Frame" and "Results of the Religious Values Framing Experiment" carefully. The researchers randomly assigned respondents to a control condition, a religious-values message without a source cue, or the same message accompanied by an evangelical source cue. The control group received neither the message nor its two follow-up questions.

The supplied file, `data/demora_2024.csv`, is a simplified teaching version of the authors' cleaned YouGov sample. Each row describes one respondent. Higher values on every outcome indicate more favorable attitudes toward refugees. Do not use `sample_weight` unless a question specifically asks you to do so.

--------------------------------------------------------------------------------
 Name                       Description
 -------------------------- -----------------------------------------------------
 `respondent_id`            Anonymous row identifier

 `treatment`                Control, religious-values message, or religious-values message with a source cue

 `refugee_thermometer`      Feeling toward refugees, from 0 to 100

 `resettlement_support`     Support for refugee resettlement, from 0 to 1

 `school_support`           Support for refugee children attending public schools, from 1 to 5

 `benefits_support`         Support for refugees receiving unemployment benefits, from 1 to 5

 `age`                      Respondent's age in 2020

 `gender`                   Respondent's gender, recorded as `Male` or `Female`

 `education`                Respondent's highest level of education

 `ideology`                 Ideology from 1, very liberal, to 5, very conservative

 `sample_weight`            YouGov weight for the target population
--------------------------------------------------------------------------------

Load the data before beginning the exercise. Beware of the relative paths; you may put the data where it suits you and edit the file path used for `read.csv`.

```{r, message=FALSE, warning=FALSE}
demora <- read.csv("data/demora_2024.csv")

demora$treatment <- factor(
  demora$treatment,
  levels = c(
    "Control",
    "Religious values message",
    "Religious values + source cue"
  )
)

summary(demora)
```

## Question 1: Research Design

1. State the paper's causal research question in your own words.
2. What are the three experimental conditions? What features distinguish them?
3. What is the unit of random assignment? What is the unit of observation in the supplied data?
4. Focus first on the religious-values message without a source cue versus the control condition. Define $T_i$, $Y_i(1)$, $Y_i(0)$, and the average treatment effect when `refugee_thermometer` is the outcome.
5. Why can we never directly observe $Y_i(1)-Y_i(0)$ for any respondent?

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]

1.State the paper’s causal research question in your own words.

연구질문 - 종교적 정체성은 당파적 선호와 다른 정책적 지향을 강화시킬 수 있는가? 
구체적으로 - 공화당원으로서의 정체성과 백인 복음주의 기독교인으로서의 정체성이 충돌할 때 어떤 일이 발생하는가?  복음주의 기독교로서 정체성은 이민정책에 대한 기존 태도를 바꾸는가? 

Theoretical questions about conflicting social identities and the potential power of counter-partisan moral messages. 

2.
The experiment contains three randomly assigned conditions ; Two Treatment groups, and one control group. The Religious Values condition introduces a moral frame emphasizing core Christian teachings . The Religious values + Source cue condition contains the exact same treatment with Religious value group but adds an explicit elite attribution stating. While control group 
Receives no text or message, with respondents moving directly to post-treatment survey question to establish a baseline measure of unprompted political attitudes. 

The experiment contains three randomly assigned conditions : religious values frame (Religious Values) , religious values frame with source cue ( Religious Values) , religious values frame with source cue(Religious Values + Source Cue) , and control. 

3.
 Unit of Random Assignment : The individual survey respondent (participant). Each individual participant was randomly assigned to one of the three experimental conditions. 

Unit of Observation: The individual respondent (individual level). Each observation/row in the dataset represents a single survey participant, containing their demographic traits, assigned treatment condition, identity strength scores, and post-treatment responses. 


4.Focus first on the religious-values message without a source cue versus the control condition. Define Ti,  Yi(1), Yi(0), and the average treatment effect when refugee_thermometer is the outcome. 

Ti is a binary indicator variable representing whether the respondent is assigned to receive the generic religious value messages or not. Ti=1 if respondent I was assigned to the Religious Values condition. Ti=0 if respondent I was assigned to the control condition. 

Yi(1) is the potential feeling thermometer score toward refugees that respondent I would express if exposed to the Religious Values Treatment (Ti=1)

Yi(0) is the potential feeling thermometer score toward refugees that responded I would express if assigned to the control condition (Ti=0) 

The Average Treatment Effect (ATE) is the expected population difference between the two potential outcomes across all individuals in the target population . 
ATE = E[Yi(1)-Yi(0)] = E[Yi(1)] - E[Yi(0)]
The Average Treatment Effect (ATE) measures the expected causal impact across the entire target population rather than just the treated group. Because random assignment makes treatment status statistically independent of potential outcomes, the observed difference between treatment and control group averages provides an unbiased estimate of the overall ATE.


5.
We cannot directly observe Yi(1) - Yi(0) because each respondent can be assigned to only one treatment condition, so we can not observe the both potential outcomes of Yi(1) and Yi(0). We call this ‘the fundamental problem of causal inference’. Random assignment allows us to overcome this fundamental problem of causal inference at the group level by using control group to estimate the average counterfactual outcome for the treated group. 






--------------------------------------------------------------------------------

## Question 2: Exploring the Experiment

1. How many observations and variables are in the supplied data?
2. How many respondents are in each experimental condition?
3. Which variables contain missing values? Show how you checked.
4. Report the mean and standard deviation of age and the proportion of respondents recorded as female.
5. Describe the distributions and scales of the four outcome variables. Explain why their numerical values should not be compared without attending to their scales.

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]

```{r}
#1  
 dim(demora) #dim
 nrow(demora)   # observations
 ncol(demora)   # variables
 
#2. 
 table(demora$treatment)
 
#3. 
 colSums(is.na(demora)) #refugee_theormometer: 25, ideology:7
 
#4.
#age 
 mean(demora$age, na.rm = TRUE) #54.96481
 sd(demora$age, na.rm = TRUE) #17.77131
#female 
 mean(demora$gender == "Female", na.rm = TRUE) #mean=0.527
 prop.table(table(demora$gender)) #0.527
 
 #5.
 # Define outcome variables
outcomes <- c(
  "refugee_thermometer",
  "resettlement_support",
  "school_support",
  "benefits_support"
)

# Summary statistics
sapply(demora[outcomes], function(x) {
  c(
    Min  = min(x, na.rm = TRUE),
    Max  = max(x, na.rm = TRUE),
    Mean = mean(x, na.rm = TRUE),
    SD   = sd(x, na.rm = TRUE)
  )
})

lapply(demora[outcomes], table, useNA = "ifany")

#their raw numerical values should not be directly compared because the same numerical difference represents different amounts of change across these scales.
 
```


--------------------------------------------------------------------------------

## Question 3: Estimating Average Treatment Effects

1. Calculate the unweighted mean of each outcome separately for the three experimental conditions.
2. For each outcome, estimate the effect of the religious-values message relative to the control using a difference in means.
3. For each outcome, estimate the effect of the religious-values message with a source cue relative to the control using a difference in means.
4. Interpret every estimate in the units of its outcome. For example, distinguish points on the 0--100 feeling thermometer from changes on the 0--1 resettlement scale.
5. Create one clearly labeled figure comparing mean `refugee_thermometer` scores across the three conditions. Preserve the experimental ordering shown in the variable table.
6. Based on these estimates, does the source cue appear to strengthen the message? Explain what comparison answers this question.

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]

```{r}
## [PUT YOUR CODE HERE]
#1. 
aggregate(
  demora[outcomes],
  by = list(treatment = demora$treatment),
  FUN = mean,
  na.rm = TRUE
)

library(dplyr)ㅇ  
demora %>%
  group_by(treatment) %>%
  summarise(
    across(
      all_of(outcomes),
      ~ mean(.x, na.rm = TRUE)
    )
  )

#2. 
sapply(outcomes, function(y) {
  mean(demora[demora$treatment == "Religious values message", y],
       na.rm = TRUE) -
  mean(demora[demora$treatment == "Control", y],
       na.rm = TRUE)
})

#3. 
sapply(outcomes, function(y) {
  mean(demora[demora$treatment == "Religious values + source cue", y],
       na.rm = TRUE) -
    mean(demora[demora$treatment == "Control", y],
         na.rm = TRUE)
})

#4.
# Religious-values message vs. Control
message_effect <- sapply(outcomes, function(y) {
  mean(demora[demora$treatment == "Religious values message", y],
       na.rm = TRUE) -
    mean(demora[demora$treatment == "Control", y],
         na.rm = TRUE)
})

# Religious-values message + source cue vs. Control
source_effect <- sapply(outcomes, function(y) {
  mean(demora[demora$treatment == "Religious values + source cue", y],
       na.rm = TRUE) -
    mean(demora[demora$treatment == "Control", y],
         na.rm = TRUE)
})

effects <- data.frame(
  Outcome = outcomes,
  Message_vs_Control = message_effect,
  SourceCue_vs_Control = source_effect
)

effects


#5.
thermo_means <- demora %>%
  group_by(treatment) %>%
  summarise(
    mean_thermometer = mean(refugee_thermometer, na.rm = TRUE),
    sd_thermometer = sd(refugee_thermometer, na.rm = TRUE),
    n = sum(!is.na(refugee_thermometer)),
    se = sd_thermometer / sqrt(n),
    lower = mean_thermometer - 1.96 * se,
    upper = mean_thermometer + 1.96 * se
  )

ggplot(
  thermo_means,
  aes(x = treatment, y = mean_thermometer)
) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    width = 0.12
  ) +
  geom_point(size = 3) +
  labs(
    title = "Mean Refugee Thermometer Score by Experimental Condition",
    x = "Experimental Condition",
    y = "Mean Refugee Thermometer Score (0–100)"
  ) +
  ylim(0, 100) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 20, hjust = 1)
  )
```
#4.
Relative to the control, the religious-values message increased the refugee thermometer by 9.36 points on the 0–100 scale and resettlement support by 0.018 units on the 0–1 scale, while decreasing school support by 0.018 points and benefits support by 0.141 points on their 1–5 scales. With the source cue, the corresponding changes were +5.32 points, +0.019 units, +0.006 points, and −0.137 points, respectively.


#6.
No. Based on the point estimates, the source cue does not appear to strengthen the message. The relevant comparison is the message-plus-source-cue condition versus the message-only condition. The estimated effects relative to control are 5.32 and 9.36 points, respectively, so the direct difference is \(5.32-9.36=-4.04\) points on the 0–100 refugee thermometer scale.

------------------------------------------------------------------------------------------

## Question 4: Randomization and Independence

1. Compare age, gender, education, and ideology across the three conditions. Present a readable balance table. You will need to explain how you summarized categorical variables.
2. Are the groups exactly identical on these observed pretreatment characteristics? Should random assignment make them exactly identical in this realized sample?
3. What does random assignment imply about the relationship between treatment assignment and both observed and unobserved pretreatment characteristics across repeated assignments?
4. Based on the research design, is $Y_i(1),Y_i(0) \perp T_i$ plausible for the religious-values-message-versus-control comparison? Explain why the design, rather than the observed balance table alone, is the basis for your answer.
5. Suppose one balance difference is statistically significant at the 0.05 level after examining twenty pretreatment variables. Would this fact alone invalidate the experiment? Explain.

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]

```{r}
#1. 
# Load packages
library(dplyr)
library(gtsummary)

# Check variables
str(demora[c("age", "gender", "education", "ideology")])

table(demora$gender, useNA = "ifany")
table(demora$education, useNA = "ifany")
table(demora$ideology, useNA = "ifany")


# Treat ideology (1-5 scale) as a categorical/ordinal variable
demora <- demora %>%
  mutate(
    ideology = factor(
      ideology,
      levels = 1:5,
      ordered = TRUE
    )
  )


# Create balance table
balance_table <- demora %>%
  select(treatment, age, gender, education, ideology) %>%
  tbl_summary(
    by = treatment,
    
    # Explicitly specify variable types
    type = list(
      age ~ "continuous",
      gender ~ "categorical",
      education ~ "categorical",
      ideology ~ "categorical"
    ),
    
    # How each variable is summarized
    statistic = list(
      age ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    
    # Number of decimal places
    digits = list(
      age ~ 1,
      all_categorical() ~ c(0, 1)
    ),
    
    missing = "ifany"
  ) %>%
  modify_header(
    label ~ "**Variable**"
  ) %>%
  bold_labels()

# Display table
balance_table

#2.

balance_table
# Age
demora %>%
  group_by(treatment) %>%
  summarise(
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE)
  )

# Gender
prop.table(
  table(demora$treatment, demora$gender),
  margin = 1
)

# Education
prop.table(
  table(demora$treatment, demora$education),
  margin = 1
)

# Ideology
prop.table(
  table(demora$treatment, demora$ideology),
  margin = 1
)

 
```
#1. 
Categorical variables (gender, education, and ideology) were summarized using the proportion of respondents in each category within each experimental condition.

#2.
No. The groups are not exactly identical on the observed pretreatment characteristics, as the balance table shows small differences in age and in the distributions of gender, education, and ideology across conditions. Random assignment does not guarantee exact balance in a particular realized sample. Rather, it produces balance in expectation across repeated random assignments, so some chance differences between groups are expected.
#3.
Across repeated random assignments, treatment assignment is independent of both observed and unobserved pretreatment characteristics. Thus, treatment groups are balanced on these characteristics in expectation, although chance imbalances may occur in any particular realized sample. 
#4.
Balance is a diagnostic check; randomization is the identification strategy.
#5. 
No. A single statistically significant balance difference would not by itself invalidate the experiment. When twenty pretreatment variables are tested at the 0.05 significance level, some significant differences can arise by chance even under proper random assignment. In fact, the probability of observing at least one significant result among twenty independent tests is approximately \(1-0.95^{20}\approx 0.64\). Therefore, an isolated significant imbalance is not sufficient evidence that randomization failed. The validity of the experiment should primarily be assessed based on the random assignment procedure rather than whether every observed covariate is statistically balanced.


---------------------------------------------------------------------------------------
## Question 5: Measurement, Assumptions, and Generalization

1. The `resettlement_support` outcome combines support for resettlement in the respondent's local community and in the United States. Discuss one advantage and one disadvantage of combining the two items.
2. The authors note that the treatment technically consists of a message plus two questions that ask respondents to reflect on it. What complication does this create when naming the causal treatment?
3. About 77 percent of treated respondents reported having heard the message previously. Does prior exposure violate random assignment? How might it affect the treatment effect estimated in this experiment?
4. State the Stable Unit Treatment Value Assumption (SUTVA) in this study. Give one plausible example of interference or treatment variation.
5. The analytic sample consists of self-identified white evangelical Republicans who passed the authors' prespecified reading-speed screen. Identify two limits this places on external validity.
6. The outcomes are survey responses measured shortly after treatment. What can and cannot be concluded about lasting changes in attitudes or political behavior?

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]
1. 지역 커뮤니티에 재정착을 지지하지 않는 사람도 미국 타 지역에서의 재정책은 지지할 수 있다.
2. 두 가지 처치의 효과 중 무엇 때문인지 정확하게 구분하기 어렵다.
3. 무작위 배정을 할 경우 문제는 해결된다. 또한 실제 효과를 보수적으로 측정하기 때문에 강건성 측면에서 의미가 있다. 
4. SUTVA 가정은 처치가 모든 대상에게 동일하고, 한 대상자에 대한 처치가 다른 대상자에 영향을 주는 간섭 효과가 있어서는 안 됨을 의미한다. 만약 실험 시행 중 참여자가 여타 교회 커뮤니티에 정보를 제시하거나 다른 실험자로부터 사전 정보를 습득했을 경우, SUTVA 가정은 위배된다. 
5. 외적 타당성은 실험 과정 및 결과가 다른 공간에서 똑같이 반복되었을 때 동일한 일반론으로 이어질 수 있는지 여부를 의미한다.
이 실험이 외적 타당성을 갖추려면 동일한 매커니즘에 관해 다른 사회적 정체성의 유권자층이 특정 의제에 대해 어떤 응답을 보이는지 관찰해야 한다.
6. 실제 실험 매커니즘에 따른 감정 변화가 선거에서의 투표 행태로까지 이어지는지 여부를 확인할 수는 없다. 

--------------------------------------------------------------------------------

##Question 6: Reading the Published Results

1. Compare your unweighted estimates with Table 2 of the paper. The authors use YouGov sample weights and regression, so your estimates need not be numerically identical. Explain what each approach targets.
2. Which outcomes provide evidence that the religious-values message increased support for refugees? Which outcomes produce null results?
3. Does the paper support the hypothesis that adding an evangelical source cue makes the religious-values message more effective? Cite the relevant comparison.
4. Summarize the paper's theoretical claim and central empirical finding in no more than 150 words.
5. Propose one follow-up experiment in a different population or political context. Clearly identify the treatment, outcome, target population, and the average treatment effect of interest.

--------------------------------------------------------------------------------

[PUT YOUR ANSWER HERE]
 #1.우리의 비가중 평균차이는 현재 분석표본에서 각 응답자를 동일하게 취급한 단순한 집단 평균 차이를 추정한다. 반면 논문의 Table 2는  YouGov 표본가중치와 회귀분석을 사용하여 표본구성의 차이를 보정한 처치효과를 추정한다. 따라서 수치는 완전히 같을 필요가 없지만, 효과의 방향과 전반적 패턴을 비교할 수 있다. 논문에서는 religious-values message가 thermometer를 약 11점, resettlement를 약 0.05 높이는 것으로 추정했다. 
 
 #2.
논문에서는 refugee thermometer에서 명확한 긍정적 효과가 나타났고, resettlement support에서도 약한 긍정적 증거가 나타났다. 반면 school support와 unemployment/benefits support에서는 긍정적 효과가 나타나지 않았다. Springer Nature Link

3. 
아니다. 이를 판단하려면 Religious Values + Source Cue 조건과 Religious Values 조건을 직접 비교해야 한다. 논문은 이 직접 비교에서 source cue가 메시지의 효과를 더 크게 만들었다는 증거를 발견하지 못했으며, 따라서 H2를 지지하지 않았다. 


3.
논문의 이론적 주장은 당파적 입장과 충돌하더라도 개인의 종교적·도덕적 가치에 호소하는 프레이밍이 정치적 태도를 변화시킬 수 있다는 것이다. 실험 결과, 친난민 기독교 가치 메시지는 백인 복음주의 공화당원의 난민에 대한 호감도와 일부 재정착 지지를 증가시켰다. 그러나 난민에 대한 정부 복지지원 태도에는 효과가 없었으며, 복음주의 출처 단서를 추가한다고 메시지 효과가 더 강해지지도 않았다. 

4. 
한국의 보수 성향 개신교인을 대상으로 실험할 수 있다. 처치는 난민 보호를 기독교적 가치와 연결한 메시지, 통제조건은 메시지가 없는 조건으로 설정한다. 결과변수는 난민 수용·재정착에 대한 지지 정도이며, 관심 있는 ATE는

--------------------------------------------------------------------------------

\newpage

# Digesting Cutting Edge Methods Papers into Key Takeaways (Building an Open-Source Reviewer Agent)

See [https://github.com/sysilviakim/quant-social-science-skills/tree/main/contrib](https://github.com/sysilviakim/quant-social-science-skills/tree/main/contrib). We are going to build an open-source reviewer agent focusing on quantitative methods. 

The TA has randomly assigned two papers to you. See `EXAMPLE-atsusaka_kim_2025.md` in the GitHub repository; your task is to carefully read the papers and create something equivalent for your assigned papers.

For PS1, choose one of the two assigned papers. For that paper, submit the following two items:

1. Show me that you actually did the reading. Print out the paper, take highlights and memos, and submit the outcome in person (I don't care if the memos and highlights are messy).
2. Submit a `.md` file that summarizes the paper's main points and explains what the reviewer should check for methodological integrity.
